How cryptic is cryptic diversity? \newline Machine learning approaches to fine scale variation in the morphology of \textit{Emys marmorata}.

### 2013-06-13 documents/turtle_presentation.Rnw
first full draft of evolution talk

  %(dissimilarity based) Evidence accumulation clustering

@@

  Here, features (principal components) predict class (subspecific assignment).

@@

\begin{frame}
  \frametitle{Model training and selection}
  Unknown appropriate number of features to ``best'' predict class. Want to minimize false positive, while maximizing true positive.

  Split data set, 75-25, training and testing.

  Tuning parameters via grid-search. Uncertainty via 10-fold cross-validation. Selection via max AUC ROC.

  Best multinomial logistic model selected via min AICc. Best random forest model via max AUC ROC.

@@

\section{Results}
\begin{frame}[fragile]
  \frametitle{Results: mophometrics}
<<gm, results = 'hide', fig.width = 3.5, fig.height = 3.5, eval = TRUE>>=
morph <- ggpairs(turtle.adult, columns = c(1:3),
                 upper = 'blank',
                 params = c(LabelSize = 2, gridLabelSize = 2, size = 1))
fits <- procGPA(turtle.land.adult)
land <- mshape.plot(fits)
morph <- putPlot(morph, land, 1, 3)
morph
@

@@

%\begin{frame}
%  \frametitle{Results: d-EAC}
%
%\end{frame}

\begin{frame}[fragile]
  \frametitle{ROC}
<<roc, eval = TRUE>>=
rf.rocs <- lapply(trf.a, function(x) {
                  rr <- x$results$ROC
                  rr})
rf.rocs <- ldply(rf.rocs)
multi.rocs <- lapply(tm.a, function(x) {
                     lapply(x, function(y) {
                            rr <- max(y$results$ROC)
                            rr})})
multi.rocs <- ldply(lapply(lapply(multi.rocs, ldply), t))

@@

\begin{frame}
  \frametitle{Generalize}
<<>>=
oo <- Reduce(cbind, Map(function(x) x$t, rr))
colnames(oo) <- groups
dd <- melt(oo)
gdist <- ggplot(dd, aes(x = value, fill = X2)) + geom_density(alpha = 0.3)
gdist <- gdist + theme(legend.title = element_blank(),
                       legend.margin = unit(0, 'cm'),
                       legend.text = element_text(size = 5),
                       axis.title = element_text(size = 8),
                       axis.text = element_text(size = 5))
gdist <- gdist + scale_fill_manual(name = '', values = cbp)

@@

\begin{frame}
  \frametitle{Best classification scheme?}
<<>>=
gg <- class.map(adult.train$spinks, adult.test$spinks, labs$spinks, 
                trf.a.analysis$class$spinks$pred, map)
gg <- gg + scale_colour_manual(name = '', values = cbp)
gg <- gg + theme(#legend.position = 'none',
                 legend.margin = unit(0, 'cm'),
                 legend.text = element_text(size = 5),
                 axis.title = element_text(size = 8),
                 axis.text = element_text(size = 5))
gg
@

\begin{frame}[fragile]
  \frametitle{Variable importance}
  % ggpairs from the best random forest model
<<imp, fig.width = 3.5, fig.height = 3.5, eval = TRUE>>=
most.imp <- trf.a$spinks$optVariables
ww <- 'spinks'
ggimp <- ggpairs(turtle.adult,
                 columns = c(most.imp[1:3], ww), 
                 colour = 'spinks',
                 upper = 'blank',
                 params = c(LabelSize = 2, gridLabelSize = 2, size = 1))

\begin{frame}
  \frametitle{Shape across most important features}

<<tps, include = FALSE>>=
fst.max <- which.max(turtle.adult[, most.imp[1]])
fst.min <- which.min(turtle.adult[, most.imp[1]])

par(mar = c(0, 0, 1, 1) + 0.1, oma = c(5, 4, 0, 0) + 0.1)

tpsgrid(turtle.land.adult[, , fst.max], mt)
tpsgrid(turtle.land.adult[, , fst.min], mt)

tpsgrid(turtle.land.adult[, , snd.max], mt)
tpsgrid(turtle.land.adult[, , snd.min], mt)

imp <- cbind(turtle.adult[, most.imp[1:2]], sp = turtle.adult$spinks)
imp.pca <- ggplot(imp, aes(x = PC3, y = PC8, colour = sp)) + geom_point()
imp.pca <- imp.pca + scale_colour_manual(name = '', values = cbp)
imp.pca
@

% show tps grids to the side of the PCA plot of the first two most important axes
\begin{figure}[h]
  \centering
  \hspace*{\fill}
  \begin{subfigure}[h]{0.3\textwidth}
    \centering
    \includegraphics[width = \textwidth]{figure/tps2}    
  \end{subfigure}
  \begin{subfigure}[h]{0.3\textwidth}
    \centering
    \includegraphics[width = \textwidth]{figure/tps5}    
  \end{subfigure}
  \begin{subfigure}[h]{0.3\textwidth}
    \centering
    \includegraphics[width = \textwidth]{figure/tps1}    
  \end{subfigure}
  \hspace*{\fill}
\end{figure}

@@

<<fig.width = 3.75, fig.height = 3.75>>=
mt <- mshape(turtle.land.adult)
# get mean of each class
spi <- turtle.adult$spinks
wspi <- lapply(levels(spi), function(x, y) which(y == x), y = spi)
mspi <- lapply(wspi, function(x, y) mshape(y[, , x]), y = turtle.land.adult)
par(mfrow = c(2, 2), mar = c(0, 0, 1, 1) + 0.1, oma = c(5, 4, 0, 0) + 0.1,
    xaxt = 'n', yaxt = 'n')
mapply(function(x, m, n) {
      tpsgrid(x, m)
      title(main = paste(n))}, x = mspi, n = levels(spi),
      MoreArgs = list(m = mt), SIMPLIFY = FALSE)
par(mfrow = c(1, 1))
@

@@

@@

\begin{frame}
  \frametitle{Acknowledgements}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
        \item Ben Frable, Dallas Krentzel, Michael Foote
        \item COLLECTIONS
        \item FUNDING AGENCIES
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \includegraphics[height = 0.25\textheight, keepaspectratio = true]{figure/chicago} \\
      \includegraphics[height = 0.25\textheight, width = \textwidth, keepaspectratio = true]{figure/field} \\
      \includegraphics[height = 0.25\textheight, keepaspectratio = true]{figure/csu}
    \end{column}
  \end{columns}
\end{frame}
\end{document}

>>>>

  Minimize sum of dissimilarities between points and medoids.

  ``Gap'' is analogous to goodness-of-clustering.

@@

  Features (principal components) predict class (subspecific assignment).

@@

\begin{frame}
  \frametitle{Model training and selection}

  \begin{itemize}
    \item split into training and testing sets, 75-25.
    \item tuning parameters via grid-search
    \item uncertainty via 10-fold CV
    \item model selection
      \begin{itemize}
        \item multinomial logistic regression: min AICc
        \item random forest: max ROC
      \end{itemize}
  \end{itemize}

@@

\section{Results}
\begin{frame}[fragile]
  \frametitle{Results: mophometrics}
<<gm, results = 'hide', fig.width = 3.5, fig.height = 3.5, eval = TRUE>>=
morph <- ggpairs(turtle.adult, columns = c(1:3),
                 upper = 'blank',
                 params = c(LabelSize = 2, gridLabelSize = 2, size = 1))
fits <- procGPA(turtle.land.adult)
land <- mshape.plot(fits)
land <- land + coord_equal()
morph <- putPlot(morph, land, 1, 3)
morph
@

@@

\begin{frame}[fragile]
  \frametitle{Model selection via ROC}
<<roc, eval = TRUE>>=
rf.rocs <- lapply(trf.a, function(x) {
                  rr <- x$results$ROC
                  rr})
rf.rocs <- ldply(rf.rocs)
multi.rocs <- lapply(tm.a, function(x) {
                     lapply(x, function(y) {
                            rr <- max(y$results$ROC)
                            rr})})
multi.rocs <- ldply(lapply(lapply(multi.rocs, ldply), t))

@@

\begin{frame}
  \frametitle{Generalize using best random forest model}
<<>>=
oo <- Reduce(cbind, Map(function(x) x$t, rr))
colnames(oo) <- groups
dd <- melt(oo)
gdist <- ggplot(dd, aes(x = value, fill = X2)) + geom_density(alpha = 0.3)
gdist <- gdist + theme(legend.title = element_blank(),
                       legend.margin = unit(0, 'cm'),
                       legend.text = element_text(size = 5),
                       axis.title = element_text(size = 8),
                       axis.text = element_text(size = 5))
gdist <- gdist + scale_fill_manual(name = '', values = cbp)

@@

\begin{frame}
  \frametitle{Best classification scheme via RF model results}
<<>>=
gg <- class.map(adult.train$spinks, adult.test$spinks, labs$spinks, 
                trf.a.analysis$class$spinks$pred, map)
gg <- gg + scale_colour_manual(name = '', values = cbp)
gg <- gg + theme(#legend.position = 'none',
                 legend.margin = unit(0, 'cm'),
                 legend.text = element_text(size = 5),
                 axis.title = element_text(size = 8),
                 axis.text = element_text(size = 5))
gg
@

@@

<<fig.width = 3.75, fig.height = 3.75>>=
# get mean of each class
spi <- turtle.adult$spinks
wspi <- lapply(levels(spi), function(x, y) which(y == x), y = spi)
mspi <- lapply(wspi, function(x, y) mshape(y[, , x]), y = turtle.land.adult)
par(mfrow = c(2, 2), mar = c(0, 0, 1, 1) + 0.1, oma = c(5, 4, 0, 0) + 0.1,
    xaxt = 'n', yaxt = 'n')
mapply(function(x, m, n) {
      tpsgrid(x, m, mag = 2)
      title(main = paste(n))}, x = mspi, n = levels(spi),
      MoreArgs = list(m = mt), SIMPLIFY = FALSE)
par(mfrow = c(1, 1))
@

@@

\begin{frame}[fragile]
  \frametitle{Variable importance of random forest model}
  % ggpairs from the best random forest model
<<imp, fig.width = 3.5, fig.height = 3.5, eval = TRUE>>=
most.imp <- trf.a$spinks$optVariables
ww <- 'spinks'
ggimp <- ggpairs(turtle.adult,
                 columns = c(most.imp[1:3], ww), 
                 colour = 'spinks',
                 upper = 'blank',
                 params = c(LabelSize = 2, gridLabelSize = 2, size = 1))

\begin{frame}
  \frametitle{Shape across PC3}

<<fig.width = 5>>=
mt <- mshape(turtle.land.adult)

fst.max <- which.max(turtle.adult[, most.imp[1]])
fst.min <- which.min(turtle.adult[, most.imp[1]])

par(mfrow = c(1, 2), mar = c(0, 0, 1, 1) + 0.1, oma = c(5, 4, 0, 0) + 0.1,
    xaxt = 'n', yaxt = 'n')
tpsgrid(turtle.land.adult[, , fst.min], mt, mag = 2)
title(main = 'min observed')
tpsgrid(turtle.land.adult[, , fst.max], mt, mag = 2)
title(main = 'max observed')
par(mfrow = c(1, 1))
@

\begin{frame}
  \frametitle{Shape across PC8}
<<fig.width = 5>>=
par(mfrow = c(1, 2), mar = c(0, 0, 1, 1) + 0.1, oma = c(5, 4, 0, 0) + 0.1,
    xaxt = 'n', yaxt = 'n')
tpsgrid(turtle.land.adult[, , snd.min], mt, mag = 2)
title(main = 'min observed')
tpsgrid(turtle.land.adult[, , snd.max], mt, mag = 2)
title(main = 'max observed')
par(mfrow = c(1, 1))
@
\end{frame}

@@

\begin{frame}
  \frametitle{Acknowledgements}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
        \item Ben Frable, Dallas Krentzel, Michael Foote, David Bapst
        \item COLLECTIONS
        \item FUNDING AGENCIES
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \includegraphics[height = 0.25\textheight, keepaspectratio = true]{figure/chicago} \\
      \includegraphics[height = 0.25\textheight, width = \textwidth, keepaspectratio = true]{figure/field} \\
      \includegraphics[height = 0.25\textheight, keepaspectratio = true]{figure/csu}
    \end{column}
  \end{columns}
\end{frame}
\end{document}
